---
title: "Technological Literacy in Developing Countries"
author: |
  | Grant Cole, Madison Lane, Noah Polacek, Luis Rodriguez, Gnandi Tanghanwaye 
  | Introduction to Data Science 
  | University of Nebraska at Omaha
date: December 15th, 2021

output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr, warn.conflicts = F)
library(ggplot2)
library(corrplot)

library(skimr)
library(psych)
library(stargazer)
library(dplyr)
library(ggplot2)
library(ggridges)
library(corrplot)
library(ggpubr)
options(dplyr.summarise.inform = FALSE)
```

\begin{abstract}
As technology becomes more widespread, people in developing countries can finally learn about these modern practices; however, not everyone learns at the same rate. After analyzing the dataset ”UNICEF Global database on Information and communications technology (ICT) skills”, we will determine if there are any characteristics about our population that yield a higher propensity of success in completing certain technological challenges. From our results, we will experiment to check if such disparities are statistically significant—we will check to see if these actually cause a variation in successfully completing technological challenges. Consequently, we aim to identify certain inequalities so that these developing countries can react accordingly.
\end{abstract}

\section{1  Introduction}

In this document, we will investigate the disparities that exist in technological development across the population of young adults in developing countries. This document utilizes data obtained directly through the United Nations Children's Fund (UNICEF) to investigate the degree to which wealth, region, and gender affect individuals' performance in completing basic tasks of technological literacy, tasks which are often taken for granted in developed nations. As there are many such variables that affect technological literacy of individuals worldwide, this topic necessitates a multifaceted investigation that utilizes various data science and statistical tools at our disposal. Additionally, we will consider various cultural factors that may influence these results. For example, many societies around the world differ greatly with their attitudes on gender inequality.   

As we seek to analyze disparities between male and female performance on technological literacy tasks, we will find that a greater understanding of the cultural differences between these societies will help us to make sense of these outcomes and put our results in context. The “tasks” that will be the basis of this investigation will include (but are not limited to) the transfer of files between computer and other devices, the use of programming languages, sending an email. Ultimately, our investigation emphasizes the role that technology has on the lives of the citizens in various developing countries.

\subsection{1.1 Project Outline}
Section 2 of this project details the process that we took to obtain and clean the data for proper analysis; more specifically, we talk about the following: cleaning of the column names, revising the original CSV file so analysis could actually be done, and altering the mode of certain data in the CSV file. Next, in Section 3, we showcase a few plots which showcase correlations between certain technological skills, and we run a few regression plots to depict such relationships. We also give some general information regarding data visualizations given later. Finally, Section 4 gives such data visualizations and gives proper explanations of these.

\subsection{1.2 Reproducability of the Document}
This entire document is prepared using R [R-base] package knitr [R-knitr]. Those who wish to reproduce and utilize the contents of this document for further analysis are entirely free to do so with proper citation.

\section{2 Description of UNICEF Technological Tasks Dataset}
As detailed earlier, the dataset used for analysis was procured from the United Nations Children Fund (UNICEF) online database. This data comes from surveying those in developing countries from early 2017 to April 2021; it describes the percentage that young adults from developing countries (males and females aged between 15-24) have completed. Furthermore, characteristics such as the following are given: the socioeconomic status, residency in a rural or urban area, and the sex of these citizens . Such technological challenges include the following: copying and pasting files, sending emails, creating computer programs, etc.
\section{3  Cleaning and Preparing the Data}
The act of cleaning the data for analysis was a momentous venture for this project. Such tasks included creating a new file for the data to be usable, mending the ids, and altering the mode of certain data in our file. Here is a list summarizing our cleaning process:

\subsection{3.1 Libraries Used to Read Dataset}  

Firstly, our dataset was fairly messy to start with. Once we imported the data, we realized that our excel file originally wasn't very compatible with RStudio—there was a lot of garbage data embedded in our set! Nevertheless, the act of cleaning our data wasn't too complicated, but our first step was to convert our file into something that we could actually utilize. To do this, we downloaded the packages readxl **[R-readxl]** and xlsx **[R-xlsx]** so our excel file could be read properly in R. These packages ended up giving us a workable dataset; however, even after we used these resources, we couldn't effectively work with the data—the ids weren't indicative of the desired data.   

\subsection{3.2 Mending Column Ids} 

Therefore, after we could actually read our data into RStudio, our next step was to fix the issues regarding the dataset's ids. Originally, many of our columns had names that were either too complicated or were incomprehensible; therefore, we ended up renaming every column and gave it a corresponding id. These columns were mostly the tasks that the population completed. Furthermore, we created a legend so these ids could be decipherable. After renaming all of our columns and assigning them a unique id, we then removed every column that had many NAs inhabited in it. This ended up bringing our rows from 64 to 48–and this gave us an optimal dataset to work with. In the following section (which is isolated to showcase its importance), we give the table that showcases the old and new ids.

\subsection{3.3 Technological Task ID Key}
While we could've potentially worked with the old ids, it would've been fairly confusing (and inefficient) to use these to work with the data. Consequently, we changed the ids of each task to the following table given in Figure 1.

```{r, echo = F, fig.allign = "center"}
library(knitr)
library(reshape2)
CleanedData <- read.csv("https://raw.githubusercontent.com/noah-po/Data-Science-Project/main/CleanedData.csv", header=TRUE)
FurtherCleanedData <- read.csv("https://raw.githubusercontent.com/noah-po/Data-Science-Project/main/FurtherCleanedData.csv")
otherCols <- c(6,7,8,9,10,11,12,13,14)
smaller <- subset(CleanedData, select = otherCols)
old_id <- data.frame(reshape2::melt(colnames(smaller),value.name = "Technological Tasks"))
CleanedData <- read.csv("https://raw.githubusercontent.com/noah-po/Data-Science-Project/main/CleanedData.csv", header=TRUE)
FurtherCleanedData <- read.csv("https://raw.githubusercontent.com/noah-po/Data-Science-Project/main/FurtherCleanedData.csv")



otherCols <- c(6,7,8,9,10,11,12,13,14)
smaller <- subset(FurtherCleanedData, select = otherCols)


newID <- data.frame(reshape2::melt((colnames(smaller)), value.name = "NewID"))

newID$oldID = old_id
kable(newID)
```
\begin{center}
Figure 1 - Technological Tasks Before and After Cleaning
\end{center}

\subsection{3.4 Changing Modes}  
Lastly, after we cleaned up the column names and gave them unique ids, we believed that we were ready to use the data; however, as we began to analyze certain variables, we ran into a fairly large issue—all numbers in the dataset were registering as strings instead of numeric values. Consequently, we couldn't calculate anything meaningful with these values, and our next step was to change the mode of columns that incorrectly described numeric values as strings. This issue was fixed by highlighting all affected columns' values in the cleaned CSV file and then changing the mode from 'general' to numeric. After this fix, our dataset was properly cleaned and ready for analysis.

\section{4  Correlation Analysis}
The dataset used for analysis contains a myriad of characteristics about our population, such as the following: the gender, geography, and socioeconomic status of young adults in these developing countries. Consequently, we felt it to be appropriate to include several data visualizations depicting these characteristics and the corresponding ability to complete technological challenges. 

\subsection{4.1 Correlation Plots} 
In this section, we explore the correlations between each characteristic and skill. Furthermore, we showcase each characteristic by isolating and checking to see if there's any correlation with completing certain tasks; below, we amalgamated these characteristics together in the following correlation plot to get the bigger picture of the state of technology in these countries.   

In figure 2 (given below), the given visualization depicts a matrix showing the various correlations (using the R values) between each of the skills in the original dataset. That is, this visual is intended to help the viewer determine which skills tend to correlate with other skills.  Skill pairs that are strongly correlated with each other include **C1** and **C8** (copying/moving folders and transferring files to different devices), **C7** and **C4** (using formulas in a spreadsheet and creating an audio/visual presentation), **C1** and **C2** (copying/moving folders and using copy-paste), and **C7** and **C3** (creating an audio visual presentation and using multimedia attachments in an email).   



```{r, echo = F}
trimmedData = select(FurtherCleanedData, -1:-5, -21,-22,-23)
correlationMat <- cor(trimmedData)
correlationMat <- correlationMat[,1:9]
corrplot(correlationMat, method = "pie", type = "upper")
```
\begin{center}
Figure 2 - Skill Correlation Plot
\end{center}

To further demonstrate the correlations mentioned above, we used graphs showing the best fit lines between those highly correlated skills, as well as their correlation coefficients:

```{r, echo = F, message = F, warning = F, out.width = "50%", }

ggplot(FurtherCleanedData, aes(C4,C7)) + geom_point() + geom_smooth(method = "lm", se=FALSE) +
  stat_regline_equation(label.y = 50, aes(label = ..rr.label..))
ggplot(FurtherCleanedData, aes(C1,C8)) + geom_point() + geom_smooth(method = "lm", se=FALSE) +
  stat_regline_equation(label.y = 50, aes(label = ..rr.label..))


```

\begin{center}
Figure 3 - Regression Line between Sending an Email and Creating an Online Presentation / between Moving a File and Transferring a File between Computers
\end{center}
```{r, echo = F, message = F, warning = F, out.width = "50%", fig.show="hold"}

ggplot(FurtherCleanedData, aes(C1,C2)) + geom_point() + geom_smooth(method = "lm", se=FALSE) +
  stat_regline_equation(label.y = 50, aes(label = ..rr.label..))
ggplot(FurtherCleanedData, aes(C3,C7)) + geom_point() + geom_smooth(method = "lm", se=FALSE) +
  stat_regline_equation(label.y = 50, aes(label = ..rr.label..))

```
\begin{center}
Figure 4 - Regression Line between Copying a File and Using a Copy and Paste Tool / Sending an Email and Creating a Presentation
\end{center}

This visualization depicts a matrix showing the various correlations (using the R values) between each of the skills in the original dataset. That is, this visual is intended to help the viewer determine which skills tend to correlate with other skills.  Skill pairs that are strongly correlated with each other include C1 and C8 (copying/moving folders and transferring files to different devices), C7 and C4 (using formulas in a spreadsheet and creating an audio/visual presentation), C1 and C2 (copying/moving folders and using copy-paste), and C7 and C3 (creating an audio visual presentation and using multimedia attachments in an email). As C1 and C8 are very strongly correlated (with an R2 value of 0.94), we can conclude that there is a strong linear correlation between possessing the ability to copy and move folders, and transfer files to different devices. Since both of these tasks are very related (as they both have to do with file transfer), it seems reasonable that individuals scoring high in one of these tasks would be likely to score highly in another.     

Similarly, C7 and C4 also share a heavy correlation (R2 = 0.94), which suggests that individuals who are more capable using spreadsheet formulas tend to score higher in their ability to create audio/visual presentations. Plainly, a casual observer may expect these results, again through similar intuition, as both of these skills share a common theme of the ability to present information to outside parties. The correlation between C1 and C2 further highlights the correlations that one would expect to see between routine technological tasks. The distinctly strong correlation (R2 = 0.9) between C7 and C3 does not seem to have the same convenient intuitive justifications as the others previously mentioned, but is certainly a notable correlation. To further demonstrate the correlations mentioned above, we used graphs showing the best fit lines between those highly correlated skills, as well as their correlation coefficients.

\subsection{4.1 Data Visualizations}
The following data visualizations showcase the general trends regarding gender, geographic, and socioeconomic factors. This is done so the reader can amalgamate and comprehend all of the information given in this project. To create these data products, we utilized the ggplot **[R-ggplot2]** package.  Furthermore, these data products may be modified for the final draft.

One of our main focuses was to analyze the genders (and corresponding region) of those who could complete these technological tasks. As a result, we believed that barcharts and histograms would be the best way to showcase the relationship between these variables.

Also, we decided to stray away from map data products in the following section. This is due to the fact that the dataset includes 48 developing countries, and there would be a lot of blank information in our map; furthermore, we can't exactly zoom in on one specific region since our dataset covers countries around the globe. We wanted to showcase the data from all of these regions in our report.

\section{5  Results}
In this section, we showcase a few data products depicting the ability of citizens in developing countries to accomplish these technological tasks. While we showcase a few data products here, we intend to add more products and analysis for our final draft. Furthermore, in the final draft, here we will give an overview of what discoveries we found in this section.

\subsection{5.1  Gender and Geographic Factors}
In this section, we focus on how gender affects one's ability to complete a certain task in these countries. We believed that combining the geographic region and the genders of the sample would be an important correlation to look out for, so we could potentially look for correlations here. Below is a data product that depicts the average percentage of people in certain regions that can complete tasks depicted in Figure 1; furthermore, the regions we're analyzing are the following: East Asia and the Pacific, Latin America and the Caribbean, North Africa and the Middle East, South America, and finally, Sub-Saharan Africa. 
```{r, echo = F}
library(dplyr)
library(ggplot2)
##Change the location of FurtherCleanedData.csv
FurtherCleanedData <- read.csv("https://raw.githubusercontent.com/noah-po/Data-Science-Project/main/FurtherCleanedData.csv")

Success.avg = FurtherCleanedData %>%group_by(Region, Sex)%>%
  summarize(Percent = mean(C1))##
ggplot(Success.avg) +  
  theme(plot.title = element_text(hjust = 0.5)) + xlab("") + geom_bar(aes(x=Sex, y=Percent, fill = Sex), stat='identity') + facet_wrap(~Region)
```
\begin{center}
Figure 5 - Percentage of people of ages 15-24 who successfully copied or moved a file or folder
\end{center}
This bar chart shows that, in this specific task, females perform better in the East Asia & Pacific and the Latin America and the Carribbean region. On the other hand, males perform drastically better in South Asia and Sub-Saharan Africa in this specific task. We suspect that this could be attributed to cultral differences between these regions; nevertheless, we're planning on analyzing this disparity more in-depth for our final draft—we hope to figure out some reasons why this observation holds. Also, this product also shows that these regions have less than a 50% success rate in this task.
```{r, echo = F}
library(dplyr)
library(ggplot2)
##Change the location of FurtherCleanedData.csv
FurtherCleanedData <- read.csv("https://raw.githubusercontent.com/noah-po/Data-Science-Project/main/FurtherCleanedData.csv")

Success.avg = FurtherCleanedData %>%group_by(Region, Sex)%>%
  summarize(Percent = mean(C2))##
ggplot(Success.avg) +   
  theme(plot.title = element_text(hjust = 0.5)) + xlab("") + geom_bar(aes(x=Sex, y=Percent, fill = Sex), stat='identity') + facet_wrap(~Region)
```
\begin{center}
Figure 6 - Percentage of people of ages 15-24 who successfully sent an email
\end{center}
This chart shows that most of the observations we made in the previous graph hold here. As we played with the data earlier, this trend seemed to be present regardless of what task we were analyzing; the percentage of successful citizens in each region were fairly uniform across each task!

\subsection{5.2  Socioeconomic Factors and Development}
In this section, we primarily focus on the impact that one's socioeconomic status has on their ability to complete technological tasks. More specifically, this section focuses on the wealth quintile of the citizens in certain regions. 
The following visualizations are all based on wealth quintiles of citizens in these developing countries. Therefore, the labels given on the visualization are in the following table:
```{r, echo = F}
otherCols <- c(18,19,20,21,22)
new_smaller <- subset(CleanedData, select = otherCols)




otherCols <- c(17,18,19,20,21)
old_id <- data.frame(reshape2::melt(colnames(new_smaller),value.name = "Technological Tasks"))

new_smaller <- subset(FurtherCleanedData, select = otherCols)

newID <- data.frame(reshape2::melt((colnames(new_smaller)), value.name = "NewID"))

newID$oldID = old_id
kable(newID)

```
\begin{center}
Figure 7 - Key of Wealth Labels
\end{center}
```{r, echo = F}
library(dplyr)

whichOnes<- c(2,17,18,19,20,21)
someClean <- subset(FurtherCleanedData, select = whichOnes)
melted <- melt(someClean, id= "Region")
melted <- melted %>%group_by(Region, variable)%>%
  summarize(Percent = mean(value))##
ggplot(melted, aes(variable, Percent), xlab="Percentages", ylab = "Region") + geom_point() + facet_wrap(~Region) +  
  theme(plot.title = element_text(hjust = 0.5)) + xlab("Wealth Quintile")


#c10 and c11
```
\begin{center}
Figure 8 - Average Percentage of People Completing Tasks Grouped by Wealth Quintile
\end{center}
This data visualization shows the ability for people in certain wealth quintiles to complete the technological tasks; the leftmost point shows the 0%-20% wealth quintile, and it increases as one goes right. For our final draft, we hope to expand upon the ideas this data product brings to light. Furthermore, this showcases the percentage of citizens that can complete at least one of these tasks—this is a different focus than the previous section.

```{r, echo = F, message = F}
library(dplyr)
whichOnes<- c(4,17,18,19,20,21)
someClean <- subset(FurtherCleanedData, select = whichOnes)
melted <- melt(someClean, id = "Development")
melted <- melted %>%group_by(Development, variable)%>%
  summarize(Percent = mean(value))##

ggplot(melted, aes(variable, Percent, fill = Development)) + geom_bar(stat = "identity") +  
  theme(plot.title = element_text(hjust = 0.5)) + xlab("Wealth Quintile") + facet_wrap(~Development)

```

```{r, echo = F, message = F}

ggplot(FurtherCleanedData, aes(C3, C6, colour = Development)) + 
geom_point() + geom_smooth(method = "lm", se=FALSE) + facet_wrap(~Sex)
```

```{r, echo = F, message = F}
ggplot(FurtherCleanedData, aes(C1, C9, colour = Development)) + 
geom_point() + geom_smooth(method = "lm", se=FALSE) + facet_wrap(~Sex)

```
\section{6  Conclusion}
In this project, we explored technological skills amongst people ages 15-24 from developing countries; furthermore, we analyzed data directly from UNICEF to look into how sex, wealth, and geographical location impacts young adults’ technological abilities. More specifically, we were able to look into the correlation between different tasks and how gender and socioeconomic factors contributed to a person’s ability to complete these tasks. Different regions had different trends in their abilities to complete different tasks. We were able to see that most regions had a large correlation between wealth and the ability to complete tasks.


\section{References}
[1]  UNICEF Database: Information communication technology skills, UNICEF, New York City, USA, https://data.unicef.org/resources/dataset/learning-and-skills/, 2021  
[2]  R Core Team, R: A Language and Environment for Statistical Computing, R Foundation for Statistical Computing, Vienna, Austria, http://www.R-project.org/, 2014  
[3]  Yihui Xieknitr:  A  general-purpose  package  for  dynamic  report  generation  in  R,http://yihui.name/knitr/, 2014
[4]  Wickham H: ggplot2: Elegant Graphics for Data Analysis. Springer-Verlag New York, https://ggplot2.tidyverse.org, 2016

